A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-05-29 sections/Metatheory.tex
further updates
English and typos

To simplify matters, we introduce an intermediate system that is still declarative
in nature, but has no overlap. This intermediate system differs only in a few 
rules from DK's declarative system; these are the changes:
\begin{enumerate}
\item Restrict the shape of $B$ in the rule $\mathtt{\forall L}$ subtyping rule:
$$
\inferrule*[right=$\mathtt{\forall L'}$]
{B \neq \all[b]B' \\ \Psi \vdash \tau \\ \Psi \vdash [\tau/a]A \le B}
{\Psi \vdash \all A \le B}
$$
\item Drop the redundant rule $\mathtt{Decl1I}$,
which can be easily derived by a combination of
$\mathtt{DeclSub}$, $\mathtt{Decl1I{\To}}$ and $\mathtt{{\le}Unit}$:
$$
\inferrule*[right=$\mathtt{DeclSub}$]
{
    \inferrule*[right=$\mathtt{Decl1I{\To}}$]
    {~}
    {\Psi \vdash () \To 1}
    \\
    \inferrule*[right=$\mathtt{{\le}Unit}$]
    {~}
    {\Psi \vdash 1 \le 1}
}
{\Psi \vdash () \Lto 1}
$$
\item Restrict the shapes of $e$ and $A$ in the subsumption rule $\mathtt{DeclSub}$:
$$
\inferrule*[right=$\mathtt{DeclSub'}$]
{e \neq \lam e' \\ A \neq \all A' \\ \Psi \vdash e \To A \\ \Psi \vdash A \le B}
{\Psi \vdash e \Lto B}
$$
\end{enumerate}
The resulting declarative system has no overlapping rules
and more closely resembles the algorithmic system,
which contains constraints of the same shape.

@@

The completeness of our algorithm means that any derivation in the
declarative system has an algorithmic counterpart.

>>>>

To simplify matters, we introduce an intermediate system that is still declarative
in nature, but has no overlap. This intermediate system differs only in a few 
rules from DK's declarative system. The changes are:
\begin{enumerate}
\item Restrict the shape of $B$ in the rule $\mathtt{\forall L}$ subtyping rule:
$$
\inferrule*[right=$\mathtt{\forall L'}$]
{B \neq \all[b]B' \\ \Psi \vdash \tau \\ \Psi \vdash [\tau/a]A \le B}
{\Psi \vdash \all A \le B}
$$
\item Drop the redundant rule $\mathtt{Decl1I}$,
which can be easily derived by a combination of
$\mathtt{DeclSub}$, $\mathtt{Decl1I{\To}}$ and $\mathtt{{\le}Unit}$:
$$
\inferrule*[right=$\mathtt{DeclSub}$]
{
    \inferrule*[right=$\mathtt{Decl1I{\To}}$]
    {~}
    {\Psi \vdash () \To 1}
    \\
    \inferrule*[right=$\mathtt{{\le}Unit}$]
    {~}
    {\Psi \vdash 1 \le 1}
}
{\Psi \vdash () \Lto 1}
$$
\item Restrict the shapes of $e$ and $A$ in the subsumption rule $\mathtt{DeclSub}$:
$$
\inferrule*[right=$\mathtt{DeclSub'}$]
{e \neq \lam e' \\ A \neq \all A' \\ \Psi \vdash e \To A \\ \Psi \vdash A \le B}
{\Psi \vdash e \Lto B}
$$
\end{enumerate}
The resulting declarative system has no overlapping rules
and more closely resembles the algorithmic system,
which contains constraints of the same shape.

@@

The completeness of our algorithm means that any derivation in the
declarative system has an algorithmic counterpart.
\jimmy{In response to "treat declarative and algorithmic worklists as a subsets":}
We explicitly relate between an algorithmic context $\Gm$ and
a declarative context $\Omega$ to avoid potential confusion.

### 2019-05-29 sections/Related.tex
English and typos

\paragraph{The Essence of ML Type-Inference} \bruno{we deal with
  higher-ranked polymorphism and do not have separate phases for 
constraint generation and solving. In such multi-phase algorithms 
there is typically a different constraint language. In contrast in our
single phase algorithm one could view the judgements as a form 
of constraints, and no separate constraint language is needed.
Other more technical differences?}

>>>>

\paragraph{The Essence of ML Type Inference}
Constraint-based type inference is adopted by \citet{remy-attapl} for
ML type systems, which do not employ higher-ranked polymorphism. An
interesting feature of their algorithm is that it keeps precise
scoping of variables, similarly to our approach.  Their algorithm is
divided into constraint generation and solving phases (which are
typical of constraint-based algorithms). Furthermore an intermediate
language is used to describe constraints and their constraint solver
utilizes a stack to track the state of the solving process.  In
contrast, our algorithm has a single phase, where the judgment chains
themselves act as constraints, thus no separate constraint language is
needed.

%A possible advantage of a separate constraint language is to provide
%a general and reuseable interface
%that interprets the desired properties for type inference.
